---
title: "zias data"
author: "Math_116"
date: "November 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tigerstats)
```

## R Markdown


```{r}
scorps <- read.csv(file="Scorpion olfaction data Nov-2016.csv",header=TRUE)
```

### check independence under marble


```{r}
marble <- scorps[which(scorps$group=="marble"), ]
x=marble$Odor.location==marble$Arm.selected
chisq.test(marble$Odor.location,marble$Arm.selected)
fisher.test(marble$Odor.location,marble$Arm.selected)
contTable <- xtabs(~ Arm.selected + Odor.location, data = marble)
barchart(contTable,
    auto.key = list(
        space = "top",
        title = "Odor.location",
        cex.title = 1,
        columns = 1),
    horizontal = FALSE)
```




## Now check for independence of those with their nose painted



Get larger class of plots

```{r}
nose <- scorps[which(scorps$group=="nosePainted"), ]
chisq.test(nose$Odor.location,nose$Arm.selected)
fisher.test(nose$Odor.location,nose$Arm.selected)
contTable <- xtabs(~ Odor.location + Arm.selected, data = nose)
barchart(contTable,
	auto.key = list(
		space = "top",
		title = "Arm.selected",
		cex.title = 1,
		columns = 1),
	horizontal = FALSE,
	xlab = "Odor Location")
```




#This shows that reaction to a marble is independent to Arm selected and also
#reaction to a predator is independent of Arm selected if your nose got painted
#comparing both is done with the Mantel_Haenszel test of common odds ratio across stratum

```{r}
conT <- xtabs(~ Arm.selected + Odor.location + group, data = scorps)
counT=as.vector(conT)
#now create the marble-nosePainted table
MNP=array(counT[5:12],dim=c(2,2,2),dimnames = list( 
Arm.selected = c("L", "R"),
          Odor.location = c("L", "R"),
          treatment = c("marble", "nosePainted")))
mantelhaen.test(x=MNP,exact=TRUE)

```
So the CMH test confirms this

##How about the untreated vs the footPainted groups

```{r}
UFP=array(c(counT[1:4],counT[13:16]),dim=c(2,2,2),dimnames = list( 
Arm.selected = c("L", "R"),
          Odor.location = c("L", "R"),
          treatment = c("untreated", "footPainted")))
mantelhaen.test(x=UFP,exact=TRUE)
```
#We see by examining the confidence interval about the common odds ratio that both groups are behaving similiarly.

## Now investigate logistic regression

```{r}
same <- scorps$Odor.location==scorps$Arm.selected
scorps$same <- same
z<-xtabs(~same +group,data=scorps)
scorps$group <- relevel(scorps$group, ref="footPainted") # Use to change the baseline
mylogit1 <- glm(same ~ group, data = scorps, family = "binomial")
scorps$group <- relevel(scorps$group, ref="nosePainted") # Use to change the baseline
mylogit2 <- glm(same ~ group, data = scorps, family = "binomial")
scorps$group <- relevel(scorps$group, ref="marble") # Use to change the baseline
mylogit3 <- glm(same ~ group, data = scorps, family = "binomial")
scorps$group <- relevel(scorps$group, ref="untreated") # Use to change the baseline
mylogit4 <- glm(same ~ group, data = scorps, family = "binomial")
summary(mylogit1)
summary(mylogit2)
summary(mylogit3)
summary(mylogit4)
```

### Making sense out of the results

So looking at mylogit1 results we have  
logit=-2.0369 +2.1322*groupmarble + 2.3470*groupnosePainted + .6506*groupuntreated

if you are in the nosePainted group then groupmarble is zero as is groupuntreated

so logit = -2.0369 + 2.347 = 0.3101 and exponentiating  exp(0.3101) we get the odds =  1.363561
which corresponds to a prob = 1.363561/(1.363561 + 1) = .5769

We can ascribe the following meaning to the individual coefficients --- they are the odds ratios between the reference group and the group in question.

for example in mylogi1 the baseline is the foot painted group so the odds ratio between the untreated and marble groups is `r summary(mylogit1)$coefficient[2]` with standard error `r  summary(mylogit1)$coefficient[2,2]`
this gives a conf interval of the odds ratio  `r summary(mylogit1)$coefficient[2] - 1.96*summary(mylogit1)$coefficient[2,2]` to   `r summary(mylogit1)$coefficient[2] + 1.96*summary(mylogit1)$coefficient[2,2]`


in sync with the CMH test

# Try contrasting prop.test with logit model


```{r}
marnose <- scorps[which((scorps$group=="marble")|(scorps$group=="nosePainted")), ]
ngroup <- factor(rep(NA,length(marnose$group)),levels=c("marble","nosePainted"))
ngroup[marnose$group %in% c("marble")] <- "marble"
ngroup[marnose$group %in% c("nosePainted")] <- "nosePainted"
same <- marnose$Odor.location==marnose$Arm.selected
marnose$same <- same
marnose$ngroup <- ngroup
z<-xtabs(~same +ngroup,data=marnose)
marnose$ngroup <- relevel(marnose$ngroup, ref="nosePainted") # Use to change the baseline
mylogit <- glm(same ~ ngroup, data = marnose, family = "binomial")
summary(mylogit)

```
# now contrast with a prop.test
```{r}
xx=table(marnose$ngroup,marnose$same)
outPTGC<- proptestGC(x=z[,2], n=(z[,1]+z[,2]))
outPTGC
```
## The prop test shows zero comfortably in the confidence interval  
if we anchor at .52 and let p1-p2 = lower.bound
then the odds ratio would be p1/(p1-.338) `r .52/(.52 - .338)` at one end  
and  p1/(p1+.338) `r .52/(.52 + .338)`

# Recant comparing untreated against marble 
## expecting independent with odds ratio of 1.

in terms of odds ratios the prop test gave
`r .52/(.52 + .338)` to `r .52/(.52 - .338)`

the CHM test gave  
95 percent confidence interval:
 0.4258014 to 5.6808653

And results from logistic regression are   
 this gives a conf interval of the odds ratio  `r summary(mylogit1)$coefficient[2] - 1.96*summary(mylogit1)$coefficient[2,2]` to   `r summary(mylogit1)$coefficient[2] + 1.96*summary(mylogit1)$coefficient[2,2]`  
 
 
# Now compare the two "AWARE" groups the Untreated and footPainted

And results from logistic regression comparing footPainted and untreated are here :    
 this gives a conf interval of the odds ratio  `r summary(mylogit1)$coefficient[4] - 1.96*summary(mylogit1)$coefficient[4,2]` to   `r summary(mylogit1)$coefficient[4] + 1.96*summary(mylogit1)$coefficient[4,2]`  
