---
title: "zias data"
author: "Math_116"
date: "November 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tigerstats)
```

## R Markdown


```{r}
scorps <- read.csv(file="Scorpion olfaction data Nov-2016.csv",header=TRUE)
```

### check independence under marble


```{r}
marble <- scorps[which(scorps$group=="marble"), ]
chisq.test(marble$Odor.location,marble$Arm.selected)
fisher.test(marble$Odor.location,marble$Arm.selected)
contTable <- xtabs(~ Arm.selected + Odor.location, data = marble)
barchart(contTable,
    auto.key = list(
        space = "top",
        title = "Odor.location",
        cex.title = 1,
        columns = 1),
    horizontal = FALSE)
```




## Now check for independence of those with their nose painted



Get larger class of plots

```{r}
nose <- scorps[which(scorps$group=="nosePainted"), ]
chisq.test(nose$Odor.location,nose$Arm.selected)
fisher.test(nose$Odor.location,nose$Arm.selected)
contTable <- xtabs(~ Odor.location + Arm.selected, data = nose)
barchart(contTable,
	auto.key = list(
		space = "top",
		title = "Arm.selected",
		cex.title = 1,
		columns = 1),
	horizontal = FALSE,
	xlab = "Odor Location")
```




#This shows that reaction to a marble is independent to Arm selected and also
#reaction to a predator is independent of Arm selected if your nose got painted
#comparing both is done with the Mantel_Haenszel test of common odds ratio across stratum

```{r}
conT <- xtabs(~ Arm.selected + Odor.location + group, data = scorps)
counT=as.vector(conT)
#now create the marble-nosePainted table
MNP=array(counT[5:12],dim=c(2,2,2),dimnames = list( 
Arm.selected = c("L", "R"),
          Odor.location = c("L", "R"),
          treatment = c("marble", "nosePainted")))
mantelhaen.test(x=MNP,exact=TRUE)

```
So the CMH test confirms this

##How about the untreated vs the footPainted groups

```{r}
UFP=array(c(counT[1:4],counT[13:16]),dim=c(2,2,2),dimnames = list( 
Arm.selected = c("L", "R"),
          Odor.location = c("L", "R"),
          treatment = c("untreated", "footPainted")))
mantelhaen.test(x=UFP,exact=TRUE)
```
#We see by examining the confidence interval about the common odds ratio that both groups are behaving similiarly.




